Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-03-20 draft_p2.tex
Updates discussion of results for GKTL

\subsubsection{Application to extreme fluctuations of the time-averaged drag}
\label{sec:gktl_drag}

@@

% Choice of parameters and perturbation
The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N_c$: $C_{gktl} = N_c \times T_a$.
In the following series of experiments we set $T_a = T = 10\tau_c$.
The sampling from the \ac{gktl} algorithm will be compared to the sampling from the control run, \textit{i.e.} a direct simulation of the flow over $T_{tot}=10^6\tau_c$.
We set $N_c=16384$, leading to $C_{gktl} \approx 1.6 \times 10^5 \tau_c$. (\textbf{Plus vraiment clair pour moi pourquoi je n'ai pas utilisé 10X plus de clones, pour matcher le coût numérique de la simulation de contrôle...})
Lastly, one must set the cloning period $\tau$.
On the one hand, too small a cloning period may result in loss of information: clones do not separate from their parent over a cloning period.
On the other hand, choosing $\tau \gg \tau_c$ may result in too few cloning stages to achieve importance sampling.
A rule of thumb is to set $\tau \approx \tau_c$.
In the following, we set $\tau = \tau_c /2$.

@@

Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf}
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F_{T_a}[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N_c}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T_{a}}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process.
Note that it corresponds to equation~\eqref{eq:estimate_biased_measure} with $k=0$.

@@

For a fixed number of trajectories $N$, there is a limited range of $k$ values for which the algorithm is able to yield an accurate estimate of $\rho_k$.
Setting $k$ to too high a value causes harmful effects resulting from the finiteness of the ensemble of trajectories: trajectories overlap over most of their duration.
This effect is illustrated in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
This is a consequence of the fact that most of the trajectories in the biased ensemble overlap, therefore leading to the same value for the averaged drag $F_T$.
See~\cite{ref1,ref2,ref3} for further details about the scaling of the required population size as $k$ is increased.

% Malgré les limitations, c'est toujours mieux
% qu'un échantillonnage direct
Unsurprisingly, sampling larger fluctuations requires a larger population size, and therefore a larger computational cost.
Even though, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to values of the time-averaged drag that could
not be sampled through a direct sampling with equal computational cost.
Table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$, for several values of $a$.
As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
This number is computed based on the assumption that the \ac{pdf} describing the random variable $F_T$ with $T=10\tau_c$ is Gaussian.
This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
While this assumption does not hold for $F_T \geq 4\sigma_T$, it may still provide numbers relevant to the comparison with the algorithm.
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
  \caption{PDF of average drag over $10\tau_c$}
  \label{fig:PDF_AVG}
\end{figure}

>>>>

\subsubsection{Application of the \ac{GKTL} algorithm to extreme fluctuations of the time-averaged drag}
\label{sec:gktl_drag}

@@

% Computational cost and subsequent choice of parameters
The computational cost $C_{gktl}$ of a \ac{gktl} experiment is determined by both the duration of the trajectories $T_a$ and the number of trajectories $N$: $C_{gktl} = N_c \times T_a$.
In the following series of experiments we set $T_a = T = 10\tau_c$.
The number of the clones is set to $N=16384$.
As a consequence, the computational cost is $C_{gktl} \approx 1.6 \times 10^5 \tau_c$.
Last but not least, one must set the cloning period $\tau$.
On the one hand, too small a cloning period may result in loss of information: clones do not separate from their parent over a cloning period.
On the other hand, choosing $\tau \gg \tau_c$ may result in too few cloning stages to achieve importance sampling.
A rule of thumb is to set $\tau \approx \tau_c$.
In the following, we set $\tau = \tau_c /2$.

@@

Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process. This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\\ref{sec:pdfs}).
Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.

@@

For a fixed number of trajectories $N$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite
size effects harm the efficiency and accuracy of the selection procedure.
Typically, at each selection step, the majority of the clones are resampled, with the expection of only one or two.
This effect is illustrated in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
This is a consequence of the fact that most of the trajectories in the biased ensemble overlap, because the resampling of
discarded trajectories is based on a very small subset of the ensemble of $N$ trajectories.
Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the
bias $k$ is increased.
An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_max$ as
a function of the ensemble size $N$. 
See~\cite{ref1,ref2,ref3} for further details.

For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the
diversity of the trajectories in the biased ensemble.
This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.

% Malgré les limitations, c'est toujours mieux
% qu'un échantillonnage direct
Sampling larger fluctuations therefore requires a larger population size, and a larger computational cost.
Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to extreme fluctuations of the time-averaged drag that could
not be sampled through a direct sampling with equal computational cost.
For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$.
Note that trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.} are counted as one.  
As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
To estimate this number we assumed that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
Even though such a Gaussian assumption does not hold for fluctuations $F_T \geq 4\sigma_T$,
it may still provide numbers relevant to the comparison with the algorithm.
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
  \caption{PDF of average drag over $10\tau_c$}
  \label{fig:PDF_AVG}
\end{figure}
